DeepSphere: a graph-based spherical CNN

### 2019-11-14 main.tex
remove trailing whitespace
indices from 1 to n for compactness

\todo{
\begin{itemize}
% 	\item shall we really name it? Or simply reference the general idea of a graph-based spherical CNN. [Yes]
% 	\item adjacency matrix: $A$ instead of $W$? nath: I am for W
% 	\item anonymize DeepSphere citation - Not doing that
% 	\item climate spherical figure: inputs with TC and AR labels superimposed [Frédérick]
	\item More details on related work? Not sure we have the space. If yes, detailed comparison with related work after method. \nati{I would drop this.}
	\item PSD of climate data: how does it compare to cosmo and 3D objects?
	\item align the spheres in figure 1
	\item mention that \citep{esteves2018sphericalcnn} and \citep{cohen2018sphericalcnn} computed equivariance error
	\item check footnotes
	\item climate data on spectrum plot \nati{I would drop this}
	\item result tables: only what supports the main message in the text, everything we have in the appendix \nati{I think this is done now}
	\item Unify vocabulary. Sampling, discretization, or pixelization of the sphere? Signal, maps, functions. Irregular, uneven. Coarsening, down-sampling.
	\item check English vs American English
	\item unify enumerations. I like (i), (ii), etc.
	\item notation: should we use indices from 1 to n instead of 0 to n-1 for compactness? \nati{I think this is already done.}
	\item add \citet{bruna2013gnn} in equivariance? They were the first to do a spherical CNN with graphs. Graph is built from a uniform random sampling of the sphere. \nati{Where do you want to add them?}
	\item figure placements \nati{Good according to me}
	\item do the experiments on the icosahedral grid with the pooling as defined in the paper \nati{I guess we are not doing that.}
	\item put all the code (learning experiments from Frédérick and equivariance experiments from Martino) in a common reproducible git repository to be linked from the paper
	\item update pygsp and merge sphere branch
	\item check that NN architectures and hyper-params are correctly reported
	\item keep NN architecture and hyper-params in the main text? Move the paragraph to appendix if not enough space. \nati{I think this does not apply anymore}
	%\item shall we still site adam and batch norm? no, mainstream enough
	\item check results for ModelNet40 and integrate in text \nati{I would drop this}
	\item Check that HEALPix is introduced in the introduction
	\item heuristic used to set the kernel width in v1. Add values to \figref{kernel_widths} for comparison.  \nati{Bit syre what you mean here...}
% 	\item check experiments section [Frédérick]: all the results numbers and NN + training details
	\item compute the invariance error of the whole NNs (before and after training, invariance can be learned) \nati{I guess we are not doing that}
	\item add it with equivariance error of the convolution to \tabref{cosmo}
	\item insert \citet{jiang2019sphericalcnn} in intro: Need a global coordinate system (ok for planets, not projections like cosmo)
\end{itemize}
}

@@

\paragraph{Theoretical convergence.}
We extend the work of \citep{belkin2005towards} to a sufficiently regular, deterministic sampling scheme. 
Following their setting, we will work with the extended graph Laplacian operator as the linear operator $L_n^t: L^{2}(\mathcal{M}) \rightarrow L^{2}(\mathcal{M})$ such that
	\begin{equation}
	\label{eq:Heat Kernel Graph Laplacian operator}
	 L_n^t f(y) := \frac{1}{n}\sum_{i=0}^{n-1} e^{ -\frac{\|x_i-y\|^2}{4t}} \left(f(y)-f(x_i)\right).
	\end{equation}
This operator extends the graph Laplacian operator previously defined with the weight scheme of \eqref{eqn:weights:belkin} to each point of the manifold, i.e. the sphere. 
As the radius of the kernel $t$ will be adapted to the number of sample, we also rescale the graph Laplacian operator as
    % \begin{equation}
	% \label{eq:rescaled Heat Kernel Graph Laplacian operator}
	 $\hat{L}_n^t := \frac{|\mathbb S^2|}{4\pi t^2}L_n^t$
	% \end{equation}
Hence, we have the following relations $\b{L}_n^t \b{f} = T_\V L_n^t f$. 
% $\b{\hat L} = \frac{|\mathbb S^2|}{4\pi t^2}\b L_n^t$. \nati{why define this, we are not using it}
Given a sampling scheme $\V = \{x_0, \dots, x_{n-1}\}$, we define $\sigma_i$ to be the patch of the surface of the sphere corresponding to $x_i$, $A_i$ its corresponding area, and $d_i$ as the largest distance between the center $x_i$ and any point on the surface $\sigma_i$.
 Define $d^{(n)} := \max_{i=0, \dots, n}d_i$ and $A^{(n)}=\max_{i=0, \dots, n}A_i$.
Given any rotation $g\in SO(3)$, we define the corresponding rotation operator $\mathcal R_g$ such that for any function $f$ on the sphere
\begin{equation}\label{eq:rotation operator}
	\mathcal R_g f(x) := f(g^{-1}(x)).
\end{equation}
\begin{theorem}
	For a sampling $\V = \{x_i\in\S^2\}_{i=0}^{n-1}$ of the sphere that is equi-area and such that $d^{(n)} \leq \frac{C}{n^\alpha}, \ \alpha\in (0,\linefrac{1}{2}]$, for all $f: \S^2 \rightarrow \R$ Lipschitz with respect to the Euclidean distance in $\R^3$, for all $y\in\S^2$, there exists a sequence $t_n = n^\beta$, $\beta\in\mathbb R$ such that 
% 	the rescaled Heat Kernel Graph Laplacian $\frac{|\S^2|}{4\pi t_n^2}\b {L}^t_n$ converges pointwise to the Laplace Beltrami operator on the sphere $\Delta{\S^2}$  
	\begin{equation*}
		\lim_{n\to\infty}\hat{L}_n^{t_n}f(y) = \triangle_{\S^2}f(y).
	\end{equation*} 
	\label{theo:pointwise convergence for a regular sampling}
\end{theorem}
As a direct implication of this theorem, the discrete counterpart of $ \mathbf{\hat L}$ converges pointwise towards the Laplace-Beltrami operator ($\Delta_{\S^2}$) in the sense explained in \citep{belkin2007convergence}. This operator commutes with rotation and hence Theorem \ref{theo:pointwise convergence for a regular sampling} is a strong step toward equivariance. We formalize in the following proposition.
%%%%%%%%%%%%%%%%%%%%%%%  link between consistency and equivariance %%%%%%%%%%%%%%%%%%%%%%%%%
\begin{theorem}\label{theo:equivariance}
Under the hypothesis of theorem \ref{theo:pointwise convergence for a regular sampling}, the rescaled graph Laplacian commutes with any rotation, in the limit of infinite sampling, i.e.,
\begin{equation*}
\forall y\in\mathbb S^2\quad\left| \mathcal R_g \hat L_n^{t_n} f (y) - \hat L_n^{t_n} \mathcal R_g f(y) \right| \xrightarrow{n\to\infty}0.
\end{equation*}
\end{theorem}
%%%%%%%%%%%%%%%%%%%%%%% end of link between consistency and equivariance %%%%%%%%%%%%%%%%%%%%%%%%%

@@

\paragraph{Preliminaries.}
The proof of theorem \ref{theo:pointwise convergence for a regular sampling} is inspired from the work of \citet{belkin2005towards}. As a result, we start by restating some of their results.
Given a sampling $\V = \{x_i\in\mathcal M\}_{i=0}^{n-1}$ of a closed, compact and infinitely differentiable manifold $\mathcal{M}$, a smooth ($\in\mathcal{C}_\infty(\mathcal{M)}$) function  $f:\mathcal{M} \rightarrow \mathbb{R}$, and defined the vector $\b{f}$ of samples of $f$ as follows: $T_\V f = \b{f} \in \mathbb{R}^n,\ \b{f}_i = f(x_i)$.
The proof is constructed by leveraging 3 different operators:
\begin{itemize}
  \item As a reminder of \eqref{eq:Heat Kernel Graph Laplacian operator}, the extended graph Laplacian operator is a linear operator $L_n^t: L^{2}(\mathcal{M}) \rightarrow L^{2}(\mathcal{M})$  is defined as
	\begin{equation}
	\label{eq:Heat Kernel Graph Laplacian operator}
	 L_n^tf(y) := \frac{1}{n}\sum_{i=0}^{n-1} e^{ -\frac{\|x_i-y\|^2}{4t}} \left(f(y)-f(x_i)\right).
	\end{equation}
	Note that we have the following relation $\b{L}_n^t \b{f} = T_\V L_n^t f$.
	\item The functional approximation to the Laplace-Beltrami operator is a linear operator $L^t: L^{2}(\mathcal{M}) \rightarrow L^{2}(\mathcal{M})$ defined  as
	\begin{equation}
    \label{eq:Functional approximation to the Laplace-Beltrami operator}
	L^tf(y) = \int_{\mathcal{M}} e^{-\frac{\|x-y\|^2}{4t}}\left(f(y)-f(x)\right)d\mu(x),
	\end{equation}
	where $\mu$ is the uniform probability measure on the manifold $\mathcal{M}$, and $\text{vol}(\mathcal{M})$ is the volume of $\mathcal{M}$.
	\item
	The Laplace-Beltrami operator $\Delta_{\mathcal{M}}$ is defined as the divergence of the gradient
	\begin{equation}
        \label{eq:laplace-beltrami}
        \Delta_{\mathcal M}f(y):= -\text{div}(\nabla_{\mathcal M}f)
    \end{equation}
    of a differentiable function $f: \mathcal M\rightarrow \R$. The gradient $\nabla f: \mathcal M \rightarrow T_p\mathcal M$ is a vector field defined on the manifold pointing towards the direction of steepest ascent of $f$, where $T_p\mathcal M$ is the affine space of all vectors tangent to $\mathcal M$ at $p$.

@@

%  \frac{|\mathbb{S}^2|}{4\pi t}
% \begin{definition}{}\\
% 	\label{def:Heat Kernel Graph Laplacian operator}
% 	\text{Given a sampling $\{x_i\in\mathcal M\}_{i=0}^{n-1}$ of the manifold we define the \textbf{operator} }$L_n^t$ such that
% 	$$L_n^tf(y) := \frac{1}{n}\left[ \sum_{i=0}^{n-1} \exp \left\{ {-\frac{||x_i-y||^2}{4t}}\right\}\left(f(y)-f(x_i)\right)\right]$$
% \end{definition}

@@

We start by proving that for smooth functions, for any fixed $t$, the extended graph Laplacian $L^t_n$ converges towards its continuous counterpart $L^t$ as the sampling increases in size.
\begin{prop}\label{prop:1}
	For an equal area sampling $\{x_i\in\S^2\}_{i=0}^{n-1}: A_i=A_j \forall i,j$ of the sphere it is true that for all $f: \S^2 \rightarrow \R$ Lipschitz with respect to the Euclidean distance $\|\cdot\|$ with Lipschitz constant $C_f$
	\begin{equation*}
	\left| \int_{\S^2}f({ x})\text{d}{\mu(x)} - \frac{1}{n}\sum_i f( x_i)\right|\leq C_f d^{(n)}.	    
	\end{equation*}
	Furthermore, for all $y\in\S^2$ the Heat Kernel Graph Laplacian operator $L^t_n$ converges pointwise to the functional approximation of the Laplace Beltrami operator $L^t$
	\begin{equation*}
	 L_n^tf(y)\xrightarrow{n\to\infty} L^tf(y).
	\end{equation*}
\end{prop}
\begin{proof}

>>>>

\todo{
\begin{itemize}
% 	\item shall we really name it? Or simply reference the general idea of a graph-based spherical CNN. [Yes]
% 	\item adjacency matrix: $A$ instead of $W$? nath: I am for W
% 	\item anonymize DeepSphere citation - Not doing that
% 	\item climate spherical figure: inputs with TC and AR labels superimposed [Frédérick]
	\item More details on related work? Not sure we have the space. If yes, detailed comparison with related work after method. \nati{I would drop this.}
	\item PSD of climate data: how does it compare to cosmo and 3D objects?
	\item align the spheres in figure 1
	\item mention that \citep{esteves2018sphericalcnn} and \citep{cohen2018sphericalcnn} computed equivariance error
	\item check footnotes
	\item climate data on spectrum plot \nati{I would drop this}
	\item result tables: only what supports the main message in the text, everything we have in the appendix \nati{I think this is done now}
	\item Unify vocabulary. Sampling, discretization, or pixelization of the sphere? Signal, maps, functions. Irregular, uneven. Coarsening, down-sampling.
	\item check English vs American English
	\item unify enumerations. I like (i), (ii), etc.
	\item add \citet{bruna2013gnn} in equivariance? They were the first to do a spherical CNN with graphs. Graph is built from a uniform random sampling of the sphere. \nati{Where do you want to add them?}
	\item figure placements \nati{Good according to me}
	\item do the experiments on the icosahedral grid with the pooling as defined in the paper \nati{I guess we are not doing that.}
	\item put all the code (learning experiments from Frédérick and equivariance experiments from Martino) in a common reproducible git repository to be linked from the paper
	\item update pygsp and merge sphere branch
	\item check that NN architectures and hyper-params are correctly reported
	\item keep NN architecture and hyper-params in the main text? Move the paragraph to appendix if not enough space. \nati{I think this does not apply anymore}
	%\item shall we still site adam and batch norm? no, mainstream enough
	\item check results for ModelNet40 and integrate in text \nati{I would drop this}
	\item Check that HEALPix is introduced in the introduction
	\item heuristic used to set the kernel width in v1. Add values to \figref{kernel_widths} for comparison.  \nati{Bit syre what you mean here...}
% 	\item check experiments section [Frédérick]: all the results numbers and NN + training details
	\item compute the invariance error of the whole NNs (before and after training, invariance can be learned) \nati{I guess we are not doing that}
	\item add it with equivariance error of the convolution to \tabref{cosmo}
	\item insert \citet{jiang2019sphericalcnn} in intro: Need a global coordinate system (ok for planets, not projections like cosmo)
\end{itemize}
}

@@

\paragraph{Theoretical convergence.}
We extend the work of \citep{belkin2005towards} to a sufficiently regular, deterministic sampling scheme.
Following their setting, we will work with the extended graph Laplacian operator as the linear operator $L_n^t: L^{2}(\mathcal{M}) \rightarrow L^{2}(\mathcal{M})$ such that
	\begin{equation}
	\label{eq:Heat Kernel Graph Laplacian operator}
	 L_n^t f(y) := \frac{1}{n}\sum_{i=1}^{n} e^{ -\frac{\|x_i-y\|^2}{4t}} \left(f(y)-f(x_i)\right).
	\end{equation}
This operator extends the graph Laplacian operator previously defined with the weight scheme of \eqref{eqn:weights:belkin} to each point of the manifold, i.e. the sphere.
As the radius of the kernel $t$ will be adapted to the number of sample, we also rescale the graph Laplacian operator as
    % \begin{equation}
	 $\hat{L}_n^t := \frac{|\mathbb S^2|}{4\pi t^2}L_n^t$
	% \end{equation}
Hence, we have the following relations $\b{L}_n^t \b{f} = T_\V L_n^t f$.
% $\b{\hat L} = \frac{|\mathbb S^2|}{4\pi t^2}\b L_n^t$. \nati{why define this, we are not using it}
Given a sampling scheme $\V = \{x_1, \dots, x_n\}$, we define $\sigma_i$ to be the patch of the surface of the sphere corresponding to $x_i$, $A_i$ its corresponding area, and $d_i$ as the largest distance between the center $x_i$ and any point on the surface $\sigma_i$.
Define $d^{(n)} := \max_{i=1, \dots, n} d_i$ and $A^{(n)} := \max_{i=1, \dots, n} A_i$.
Given any rotation $g\in SO(3)$, we define the corresponding rotation operator $\mathcal R_g$ such that for any function $f$ on the sphere
\begin{equation}\label{eq:rotation operator}
	\mathcal R_g f(x) := f(g^{-1}(x)).
\end{equation}
\begin{theorem}
	For a sampling $\V = \{x_i\in\S^2\}_{i=1}^n$ of the sphere that is equi-area and such that $d^{(n)} \leq \frac{C}{n^\alpha}, \ \alpha\in (0,\linefrac{1}{2}]$, for all $f: \S^2 \rightarrow \R$ Lipschitz with respect to the Euclidean distance in $\R^3$, for all $y\in\S^2$, there exists a sequence $t_n = n^\beta$, $\beta\in\mathbb R$ such that
% 	the rescaled Heat Kernel Graph Laplacian $\frac{|\S^2|}{4\pi t_n^2}\b {L}^t_n$ converges pointwise to the Laplace Beltrami operator on the sphere $\Delta{\S^2}$
	\begin{equation*}
		\lim_{n\to\infty}\hat{L}_n^{t_n}f(y) = \triangle_{\S^2}f(y).
	\end{equation*}
	\label{theo:pointwise convergence for a regular sampling}
\end{theorem}
As a direct implication of this theorem, the discrete counterpart of $ \mathbf{\hat L}$ converges pointwise towards the Laplace-Beltrami operator ($\Delta_{\S^2}$) in the sense explained in \citep{belkin2007convergence}. This operator commutes with rotation and hence Theorem \ref{theo:pointwise convergence for a regular sampling} is a strong step toward equivariance. We formalize in the following proposition.
%%%%%%%%%%%%%%%%%%%%%%%  link between consistency and equivariance %%%%%%%%%%%%%%%%%%%%%%%%%
\begin{theorem}\label{theo:equivariance}
Under the hypothesis of theorem \ref{theo:pointwise convergence for a regular sampling}, the rescaled graph Laplacian commutes with any rotation, in the limit of infinite sampling, i.e.,
\begin{equation*}
\forall y\in\mathbb S^2\quad\left| \mathcal R_g \hat L_n^{t_n} f (y) - \hat L_n^{t_n} \mathcal R_g f(y) \right| \xrightarrow{n\to\infty}0.
\end{equation*}
\end{theorem}
%%%%%%%%%%%%%%%%%%%%%%% end of link between consistency and equivariance %%%%%%%%%%%%%%%%%%%%%%%%%

@@

\paragraph{Preliminaries.}
The proof of theorem \ref{theo:pointwise convergence for a regular sampling} is inspired from the work of \citet{belkin2005towards}. As a result, we start by restating some of their results.
Given a sampling $\V = \{x_i\in\mathcal M\}_{i=1}^n$ of a closed, compact and infinitely differentiable manifold $\mathcal{M}$, a smooth ($\in\mathcal{C}_\infty(\mathcal{M)}$) function  $f:\mathcal{M} \rightarrow \mathbb{R}$, and defined the vector $\b{f}$ of samples of $f$ as follows: $T_\V f = \b{f} \in \mathbb{R}^n,\ \b{f}_i = f(x_i)$.
The proof is constructed by leveraging 3 different operators:
\begin{itemize}
  \item As a reminder of \eqref{eq:Heat Kernel Graph Laplacian operator}, the extended graph Laplacian operator is a linear operator $L_n^t: L^{2}(\mathcal{M}) \rightarrow L^{2}(\mathcal{M})$ defined as
	\begin{equation}
%	\label{eq:Heat Kernel Graph Laplacian operator}
	 L_n^t f(y) := \frac{1}{n}\sum_{i=1}^{n} e^{ -\frac{\|x_i-y\|^2}{4t}} \left(f(y)-f(x_i)\right).
	\end{equation}
	Note that we have the following relation $\b{L}_n^t \b{f} = T_\V L_n^t f$.
	\item The functional approximation to the Laplace-Beltrami operator is a linear operator $L^t: L^{2}(\mathcal{M}) \rightarrow L^{2}(\mathcal{M})$ defined  as
	\begin{equation}
    \label{eq:Functional approximation to the Laplace-Beltrami operator}
	L^tf(y) = \int_{\mathcal{M}} e^{-\frac{\|x-y\|^2}{4t}}\left(f(y)-f(x)\right)d\mu(x),
	\end{equation}
	where $\mu$ is the uniform probability measure on the manifold $\mathcal{M}$, and $\text{vol}(\mathcal{M})$ is the volume of $\mathcal{M}$.
	\item
	The Laplace-Beltrami operator $\Delta_{\mathcal{M}}$ is defined as the divergence of the gradient
	\begin{equation}
        \label{eq:laplace-beltrami}
        \Delta_{\mathcal M}f(y) := -\text{div}(\nabla_{\mathcal M}f)
    \end{equation}
    of a differentiable function $f: \mathcal M\rightarrow \R$. The gradient $\nabla f: \mathcal M \rightarrow T_p\mathcal M$ is a vector field defined on the manifold pointing towards the direction of steepest ascent of $f$, where $T_p\mathcal M$ is the affine space of all vectors tangent to $\mathcal M$ at $p$.

@@

%  \frac{|\mathbb{S}^2|}{4\pi t}
% \begin{definition}{}\\
% 	\label{def:Heat Kernel Graph Laplacian operator}
% 	\text{Given a sampling $\{x_i\in\mathcal M\}_{i=1}^n$ of the manifold we define the \textbf{operator} }$L_n^t$ such that
% 	$$L_n^tf(y) := \frac{1}{n}\left[ \sum_{i=1}^n \exp \left\{ {-\frac{||x_i-y||^2}{4t}}\right\}\left(f(y)-f(x_i)\right)\right]$$
% \end{definition}

@@

We start by proving that for smooth functions, for any fixed $t$, the extended graph Laplacian $L^t_n$ converges towards its continuous counterpart $L^t$ as the sampling increases in size.
\begin{prop}\label{prop:1}
	For an equal area sampling $\{x_i\in\S^2\}_{i=1}^n: A_i=A_j \forall i,j$ of the sphere it is true that for all $f: \S^2 \rightarrow \R$ Lipschitz with respect to the Euclidean distance $\|\cdot\|$ with Lipschitz constant $C_f$
	\begin{equation*}
	\left| \int_{\S^2}f({ x})\text{d}{\mu(x)} - \frac{1}{n}\sum_i f( x_i)\right|\leq C_f d^{(n)}.
	\end{equation*}
	Furthermore, for all $y\in\S^2$ the Heat Kernel Graph Laplacian operator $L^t_n$ converges pointwise to the functional approximation of the Laplace Beltrami operator $L^t$
	\begin{equation*}
	 L_n^tf(y)\xrightarrow{n\to\infty} L^tf(y).
	\end{equation*}
\end{prop}
\begin{proof}
